A User-Extensible and Adaptable Parser Architecture

### 2008-08-22 doc/sgai-2008.tex
Changes suggested by the diction program.

\abstract{%
    Some parsers need to be very precise and strict in what they parse, yet
    must allow users to easily adapt or extend the parser to parse new
    inputs, without requiring that the user have an in-depth knowledge and
    understanding of the parser's internal workings.  This paper presents a
    novel parsing architecture, designed for parsing Postfix log files,
    which aims to make the process of parsing new inputs as simple as
    possible, enabling users to trivially add new rules (to parse variants
    of existing inputs) and relatively easily add new actions (to process a
    previously unknown class of input).  The architecture scales linearly
    as the number of rules and/or size of input increases, making it
    suitable for parsing large corpora or months of accumulated data.
}

@@

The architecture described herein was developed as part of a larger project
to improve anti-spam techniques by analysing the performance of the set of
filters currently in use, optimising the order and membership of the set
based on that analysis, and developing supplemental filters where
deficiencies are identified.  Most anti-spam techniques are
content-based~\cite{a-plan-for-spam, word-stemming, relaxed-online-svms}
which requires accepting the mail in the first place to determine if it is
spam, but rejecting mail during the delivery attempt is preferable: senders
of non-spam mail that is mistakenly rejected will receive an immediate
non-delivery notice; load is reduced on the accepting mail server (allowing
more intensive content-based techniques to be used); users have less mail
to wade through.  Improving the performance of anti-spam techniques which
are applied when mail is being transferred via \SMTP{}\footnotemark{} is
the goal of this project, by providing a platform for reasoning about
anti-spam filters.  The approach chosen to measure performance is to
analyse the log files produced by the \SMTP{} server in use,
Postfix~\cite{postfix}, rather than modifying it to generate statistics:
this approach improves the chances of other sites testing and utilising the
software.  The need arose for a parser capable of dealing with the great
number and variety of log lines produced by Postfix: the parser must be
designed so that adding support for parsing new inputs is a simple task, as
the log lines to be parsed will change over time.  The variety in log lines
occurs for several reasons:\footnotetext{Simple Mail Transfer Protocol
transfers mail across the Internet from the sender to one or more
recipients.  It is a simple, human readable, plain text protocol, making it
quite easy to test and debug problems with it.  The original protocol
definition is RFC~821~\cite{RFC821}, updated in RFC~2821~\cite{RFC2821}.}

@@

The solution developed is conceptually simple: provide a few generic
functions (\textit{actions\/}), each capable of dealing with an entire
category of log lines (e.g.\ rejecting a mail), accompanied by a multitude
of precise patterns (\textit{rules\/}), each of which matches all log lines
of a specific type and only that type (e.g.\ rejection by a specific
\DNSBL{}).  It is an accepted standard to separate the parsing procedure
from the declarative grammar it operates with; part of the novelty here is
in the way that the grammar is itself partially procedural (each action is
a separate procedure).  This architecture is ideally suited to parsing
inputs where the input is not fully understood or does not conform to a
fixed grammar: the architecture warns about unparsed inputs and other
errors, but continues on parsing as best it can, allowing the developer of
a new parser to decide which deficiencies are most important and require
attention first, rather than being forced to fix the first error which
arises.

@@

    \item [Framework]  The framework is the structure which actions and
        rules plug into.  It provides the parsing loop, loading and
        validation of rules, shared data storage, storage of results, and
        other support functions.

@@

    \item [Rules]  The rules are responsible for classifying or matching
        inputs, specifying the action to invoke and the regex which matches
        the inputs and extracts data.

@@

For each input the framework tries each rule in turn until it finds a rule
which matches the input, then invokes the action specified by that rule.

@@

There is some similarity (shown in \tableref{Similarities with ATN})
between the architecture and William Wood's \ATN{}~\cite{atns, nlpip}, a
tool used in Computational Linguistics for creating grammars to parse or
generate sentences.  The resemblance between \ATN{} and the architecture is
accidental, but it is interesting how two different approaches share a
similar division of responsibilities, while having different semantics.

@@

\begin{table}[hbtp]
    \caption{Similarities with ATN}\label{Similarities with ATN}
    % The \smallskip{} below stop the text hitting the lines.
    \begin{tabular}[]{lll}
        \tabletopline{}%
        ATN           & Parser Architecture & Similarity                                \\
        \tablemiddleline{}%
        Networks      & Framework           & Determines the sequence of transitions or \\
                      &                     & actions which constitutes a valid input.  \\
        Transitions   & Actions             & Saves data and imposes conditions the     \\
                      &                     & input must meet to be considered valid.   \\
        Abbreviations & Rules               & Responsible for classifying input.        \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

@@

The framework takes care of miscellaneous support functions and low level
details of parsing, freeing the authors of actions to concentrate on
writing productive code.  It provides the link between actions and rules,
allowing either to be improved independently of the other.  It provides
shared storage to pass data between actions, loads and validates rules,
manages parsing, invokes actions, tracks how often each rule matches to
optimise rule ordering (\sectionref{Rule ordering}), and stores results in
the database.  The framework is largely independent of the parser being
written: the same functionality will be required for most or all parsers,
though some specialised support functions may be provided.  The framework
is the core of the architecture and is deliberately quite simple: the rules
deal with the variation in inputs, and the actions deal with the
intricacies and complications encountered when parsing.

The function which finds the rule matching the input and invokes the
requested action can be expressed in pseudo-code as:

@@

Each action contains the code required to deal with a particular category
of input, e.g.\ rejections.  The actions are parser-specific: each parser
author will need to write the required actions from scratch, unless
extending an existing parser.  It is expected that parsers utilising this
architecture will have a high ratio of rules to actions, hopefully
resulting in simpler rules and clearer distinctions between the inputs
parsed by different rules.  In the Postfix log parser developed for this
project there are \numberOFactions{} actions and \numberOFrules{} rules,
with a very uneven distribution of rules to actions as shown in
\graphref{Distribution of rules per action}.
\showgraph{build/graph-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} Unsurprisingly the action with
the highest number of associated rules is \texttt{DELIVERY\_REJECTED}, the
action which handles Postfix rejecting a mail delivery attempt; it is
followed by \texttt{SAVE\_DATA}, the action responsible for handling
informative log lines, which supplement the data gathered from other log
lines but do not by themselves represent interesting events.  The third
most common action is, perhaps surprisingly, \texttt{UNINTERESTING}: this
action does nothing when executed, allowing uninteresting log lines to be
parsed without causing any effects (it does not imply that the input is
ungrammatical or unparsed).  Generally rules specifying the
\texttt{UNINTERESTING} action are rules to parse log lines which are not
associated with a specific mail, e.g.\ notices about configuration files
changing.  The remaining actions only have one or two associated rules: in
some cases this is because there will only ever be one log line variant for
that action, e.g.\ all log lines showing that a remote client has connected
are matched by a single rule and handled by the \texttt{CONNECT} action; in
other cases the action is required to address a deficiency in the log
files, or a complication which arises during parsing.  Taking the
\texttt{CONNECT} action as an example, it creates a new data structure in
memory for each client connecting, and saves the data extracted by the
rule; this data will subsequently be entered into the database.  If a data
structure already exists for that connection it is treated as a symptom of
a bug, and the action issues a warning containing the full contents of the
existing data structure, plus the log line which has just been parsed.

The ability to add special purpose actions to deal with difficulties and
new requirements which are discovered during parser development is one of
the strengths of this architecture.  Instead of writing one monolithic
function which must be modified to support new behaviour, with all the
attendant risks of adversely affecting the existing parser, when a new
requirement arises an independent action can be written to satisfy it.  In
some cases the new action will require the cooperation of other actions,
e.g.\ to set or check a flag.  There is a possibility of introducing
failure when modifying existing actions, but the modifications will be
smaller and occur less frequently than with a monolithic architecture, thus
failures will be less likely and will be easier to test for and diagnose.
The architecture can be implemented in an object oriented style, allowing
sub-classes to extend or override actions in addition to adding new
actions; because each action is an independent procedure, the sub-class
need only modify the action it is overriding, rather than reproducing large
chunks of functionality.

During development of the Postfix log parser it became apparent that in
addition to the obvious variety in log lines, there were a large number of
complications to be overcome.  Some were the result of deficiencies in
Postfix's logging (some of which were improved on or resolved by later
versions of Postfix); others were due to the vagaries of process
scheduling, client behaviour, and administrative actions.  All were
successfully accommodated in the Postfix log parser: adding new actions was
enough to overcome several of the complications; others required
modifications to a single existing action to work around the difficulties;
the remainder were resolved by adapting existing actions to cooperate and
exchange extra data, changing their behaviour as appropriate.

Actions may return a modified input line which will be parsed as if read
from the input stream, allowing for a simplified version of cascaded
parsing~\cite{cascaded-parsing}.  This is a powerful facility allowing a
number of rules and actions to parse a single input.

@@

Rules categorise inputs, specifying both the regex to match against each
input and the action to invoke when the match is successful.  The Postfix
log parser stores its rules in the same SQL database the results are stored
in, removing any doubt as to which set of rules was used to produce a set
of results; other implementations are free to store the rules in whatever
fashion suits their needs.  The framework warns about every unparsed input,
to alert the user that they need to alter or extend their ruleset; the
Postfix log parser successfully parses every log line in the
\numberOFlogFILESall{} log files it is currently being tested with.  The
framework requires each rule to have \texttt{action} and \texttt{regex}
attributes; each implementation is free to add any additional attributes it
requires.  The Postfix log parser adds several attributes to optimise rule
ordering (\sectionref{Rule ordering}), to restrict which log lines each
rule can be matched against, and to describe each rule.  Data is
automatically extracted and saved from the log line based on keywords in
the rule's regex, but there is also a facility to specify extra data to be
saved.

Parsing new inputs is generally achieved by creating a new rule that pairs
an existing action with a new regex.  Adding new rules is simple: define
all the required attributes and append the new rule to wherever the
existing rules are stored.  The Postfix log parser supplies a utility based
on Simple Logfile Clustering Tool~\cite{slct-paper} to aid in producing
regexes from previously unparsed log lines.  Other approaches to rule
management can be used, e.g.\ instead of manually adding new rules, machine
learning techniques could be used to automatically generate new rules,
based on the existing rules and input lines.  If this approach was taken
the choice of machine learning technique would be constrained by the size
of typical data sets (see \sectionref{Results}).  Techniques which require
the full data set when training would be impractical: Instance Based
Learning~\cite{instance-based-learning} techniques which automatically
determine which inputs from the training set are valuable and which inputs
can be discarded would probably reduce the data required to a manageable
size.  This avenue of research and development has not been pursued by the
authors, but could easily be undertaken independently.

The architecture does not try to detect overlapping rules: that
responsibility is left to the author of the rules.  Unintentionally
overlapping rules lead to inconsistent parsing and data extraction because
the first matching rule wins, and the order in which rules are tried
against each input may change between parser invocations.  Overlapping
rules are frequently a requirement, allowing a more specific rule to match
some inputs and a more general rule to match the remainder, e.g.\
separating \SMTP{} delivery to specific sites from \SMTP{} delivery to the
rest of the world.  This can simplify both the general rule and the more
specific rule; additionally rules from different sources can be combined
with a minimum of prior cooperation or modification required.  Overlapping
rules should have a priority attribute to specify their relative ordering;
negative priorities may be useful for catchall rules.

Storing the rules separately to the parser allows external tools to be
written to detect overlapping rules.  Traditional regexes are equivalent in
computational power to \FA{}, and it is possible to convert between the
two, so regex overlap could be detected by finding a non-empty intersection
of two \FA{}\@.  The standard equation for \FA{} intersection (given for
example in~\cite{intersection-of-NFA-using-Z}) is: $FA1 \cap{} FA2 =
\overline{(\overline{FA1} \cup{} \overline{FA2})}$.  Perl 5.10 regexes are
more powerful than traditional regexes, e.g.\ it is possible to match
correctly balanced brackets nested to an arbitrary depth:
\newline{}\tab{}\verb!/^[^<>]*(<(?:(?>[^<>]+)|(?1))*>)[^<>]*$/!\newline{}
Perl 5.10 regexes can maintain an arbitrary state stack and are thus
equivalent in computational power to \PDA{} or \CFL{}, so detecting overlap
may require calculating the intersection of two \PDA{} or \CFL{}\@.  The
intersection of two \CFL{} is not closed, i.e.\ the resulting language
cannot in all cases be parsed by a \CFL{}, so intersection may be
intractable in some cases e.g.:
$a^{*}b^{n}c^{n}~\cap~a^{n}b^{n}c^{*}~\rightarrow~a^{n}b^{n}c^{n}$.

@@

It is possible to define pathological regexes, which fall into two main
categories: regexes which match every input, and regexes which consume
excessive amounts of CPU time during matching.  Defining a regex which
matches every input is trivial: \verb!/^/! matches the start of every
input.  In general a regex will not consume much CPU time when the match is
successful; usually CPU time is consumed when a regex with a lot of
alteration and variable quantifiers fails to match.  This topic is beyond
the scope of this paper, see~\cite{mastering-regular-expressions} for
in-depth discussion.

@@

    \item [Partially context-sensitive actions and cascaded parsing:]
        Actions can consult the results (or lack of results) of previous
        actions during execution, providing some context sensitivity.
        Actions can not inspect past or future inputs, however actions can
        return a modified input that will be parsed as if read from the
        input stream, allowing for a simplified version of cascaded
        parsing~\cite{cascaded-parsing}.

@@

    \item [Closer to Natural Language Processing than using a fixed
        grammar:] Unlike traditional parsers such as those used when
        compiling a programming language, this architecture does not
        require a fixed grammar specification which inputs must adhere to.
        The architecture is capable of dealing with interleaved inputs, out
        of order inputs, and ambiguous inputs where heuristics must be
        applied --- all have arisen and been successfully accommodated in
        the Postfix log parser.

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important for this architecture because it is
expected that typical parsers will have a large number of rules.  There are
\numberOFrules{} rules in the full Postfix log parser ruleset, whereas the
minimum number of rules required to parse the \numberOFlogFILES{} log files
is \numberOFrulesMINIMUM{}, \numberOFrulesMINIMUMpercentage{} of the full
ruleset.  A second set of statistics was generated using the minimum
ruleset, and the parsing times compared to those generated using the full
ruleset: the percentage parsing time increase when using the full ruleset
instead of the minimal ruleset for optimal, shuffled and reversed orderings
is shown in \graphref{Percentage parsing time increase of maximum ruleset
over minimum ruleset}.  It is clear that the increased number of rules has
a noticeable performance impact with reverse ordering, and a lesser impact
with shuffled ordering.  The optimal ordering shows a mean increase of
\input{build/include-full-ruleset-vs-minimum-ruleset} in parsing time for a
\numberOFrulesMAXIMUMpercentage{} increase in the number of rules.  These
results show that the architecture scales very well as the number of rules
increases, and that optimally ordering the rules is an important
optimisation enabling this.
\showgraph{build/graph-full-ruleset-vs-minimum-ruleset}{Percentage parsing
time increase of maximum ruleset over minimum ruleset}{Percentage parsing
time increase of maximum ruleset over minimum ruleset}

@@

Correct parsing of log lines must be measured first.  Warnings are issued
for any log lines which are not parsed; no such warnings are issued while
parsing the \numberOFlogFILES{} log files, so it can be safely concluded
that there are zero false negatives.  False positives are harder to
quantify: manually verifying that the correct rule parsed each of the
\numberOFlogLINES{} log lines is infeasible.  A random sample of 6,039 log
lines (0.00994\% of \numberOFlogLINES{}) was parsed and the results
manually verified by inspection to ensure that the correct rule parsed each
log line.  The sample results contained zero false positives, and this
check has been automated to ensure continued accuracy.  The authors are
confident that zero false positives occur when parsing the
\numberOFlogFILES{} log files.

The proportion of mail delivery attempts correctly understood is much more
difficult to determine accurately than the proportion of log lines
correctly parsed.  The implementation can dump its state tables in a human
readable form; examining these tables with reference to the log files and
database is the best way to detect mails which were not handled properly.
The implementation issues warnings when it detects any errors or
discrepancies, which should alert the user to a problem.  There should be
few or no warnings during parsing, and when parsing is finished the state
table should only contain entries for mails starting before or ending after
the log file.  A second sample of 6000 log lines was parsed with all
debugging options enabled, resulting in 167,448 lines of output.  All
167,448 lines were examined in conjunction with the log segment and a dump
of the resulting database, verifying that for each of the log lines the
implementation performed correctly.  The implementation produced 4 warnings
about deficiencies in the log segment, 10 mails correctly remaining in the
state tables, and 1625 correct entries in the database.  No error messages
were produced, therefore there were no false negatives.  Given the evidence
detailed above, the authors are confident that the false positive rate when
understanding a mail delivery attempt is zero.

@@

This architecture's greatest strength is the ease with which it can be
adapted to deal with new requirements and inputs.  Parsing a variation of
an existing input is a trivial task: simply modify an existing rule or add
a new rule with an appropriate regex and the task is complete.  Parsing a
new category of input is achieved by writing a new action and appropriate
rules; quite often the new action will not need to interact with existing
actions, but when interaction is required the framework provides the
necessary facilities.  The decoupling of rules from actions allows
different sets of rules to be used with the same actions, e.g.\ a parser
might have actions to process versions one and two of a file format; the
parser will then parse version one, or version two, or both versions,
depending on the ruleset selected.  The architecture makes it possible to
apply commonly used programming techniques such as object orientation,
inheritance, composition, delegation, and roles when designing and
implementing a parser, simplifying the process of working within a team
with shared responsibility for a parser, or when developing and testing
additional functionality.  This architecture is ideally suited to parsing
inputs where the input is not fully understood or does not follow a fixed
grammar: the architecture warns about unparsed inputs and other errors, but
continues on parsing as best it can, allowing the developer of a new parser
to decide which deficiencies are most important and require attention
first, rather than being forced to fix the first error which arises.

>>>>

\abstract{%
    Some parsers need to be very precise and strict in what they parse, yet
    must allow users to easily adapt or extend the parser to parse new
    inputs, without requiring that the user have an in-depth knowledge and
    understanding of the parser's internal workings.  This paper presents a
    novel parsing architecture, designed for parsing Postfix log files,
    that aims to make the process of parsing new inputs as simple as
    possible, enabling users to trivially add new rules (to parse variants
    of existing inputs) and relatively easily add new actions (to process a
    previously unknown class of input).  The architecture scales linearly
    as the number of rules and/or size of input increases, making it
    suitable for parsing large corpora or months of accumulated data.
}

@@

The architecture described herein was developed as part of a larger project
to improve anti-spam techniques by analysing the performance of the set of
filters currently in use, optimising the order and membership of the set
based on that analysis, and developing supplemental filters where
deficiencies are identified.  Most anti-spam techniques are
content-based~\cite{a-plan-for-spam, word-stemming, relaxed-online-svms}
which requires accepting the mail in the first place to determine if it is
spam, but rejecting mail during the delivery attempt is preferable: senders
of non-spam mail that is mistakenly rejected will receive an immediate
non-delivery notice; load is reduced on the accepting mail server (allowing
more intensive content-based techniques to be used); users have less mail
to wade through.  Improving the performance of anti-spam techniques that
are applied when mail is being transferred via \SMTP{}\footnotemark{} is
the goal of this project, by providing a platform for reasoning about
anti-spam filters.  The approach chosen to measure performance is to
analyse the log files produced by the \SMTP{} server in use,
Postfix~\cite{postfix}, rather than modifying it to generate statistics:
this approach improves the chances of other sites testing and utilising the
software.  The need arose for a parser capable of dealing with the great
number and variety of log lines produced by Postfix: the parser must be
designed so that adding support for parsing new inputs is a simple task, as
the log lines to be parsed will change over time.  The variety in log lines
occurs for several reasons:\footnotetext{Simple Mail Transfer Protocol
transfers mail across the Internet from the sender to one or more
recipients.  It is a simple, human readable, plain text protocol, making it
quite easy to test and debug problems with it.  The original protocol
definition is RFC~821~\cite{RFC821}, updated in RFC~2821~\cite{RFC2821}.}

@@

The solution developed is conceptually simple: provide a few generic
functions (\textit{actions\/}), each capable of dealing with an entire
category of log lines (e.g.\ rejecting a mail), accompanied by a multitude
of precise patterns (\textit{rules\/}), each of which matches all log lines
of a specific type and only that type (e.g.\ rejection by a specific
\DNSBL{}).  It is an accepted standard to separate the parsing procedure
from the declarative grammar it operates with; part of the novelty here is
in the way that the grammar is itself partially procedural (each action is
a separate procedure).  This architecture is ideally suited to parsing
inputs where the input is not fully understood or does not conform to a
fixed grammar: the architecture warns about unparsed inputs and other
errors, but continues on parsing as best it can, allowing the developer of
a new parser to decide which deficiencies are most important and require
attention first, rather than being forced to fix the first error that
arises.

@@

    \item [Framework]  The framework is the structure that actions and
        rules plug into.  It provides the parsing loop, shared data
        storage, loading and validation of rules, storage of results, and
        other support functions.

@@

    \item [Rules]  The rules are responsible for classifying or matching
        inputs, specifying the action to invoke and the regex that matches
        the inputs and extracts data.

@@

For each input the framework tries each rule in turn until it finds a rule
that matches the input, then invokes the action specified by that rule.

@@

Some similarity exists between the architecture and William Wood's
\ATN{}~\cite{atns, nlpip}, a tool used in Computational Linguistics for
creating grammars to parse or generate sentences.  The resemblance between
\ATN{} and the architecture (shown in \tableref{Similarities with ATN}) is
accidental, but it is obvious that the two approaches share a similar
division of responsibilities, although having different semantics.

@@

\begin{table}[hbtp]
    \caption{Similarities with ATN}\label{Similarities with ATN}
    % The \smallskip{} below stop the text hitting the lines.
    \begin{tabular}[]{lll}
        \tabletopline{}%
        ATN           & Parser Architecture & Similarity                                \\
        \tablemiddleline{}%
        Networks      & Framework           & Determines the sequence of transitions or \\
                      &                     & actions that constitutes a valid input.   \\
        Transitions   & Actions             & Saves data and imposes conditions the     \\
                      &                     & input must meet to be accepted as valid.  \\
        Abbreviations & Rules               & Responsible for classifying input.        \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

@@

The framework takes care of miscellaneous support functions and low level
details of parsing, freeing the programmers writing actions to concentrate
on writing productive code.  It provides the link between actions and
rules, allowing either to be improved independently of the other.  It
provides shared storage to pass data between actions, loads and validates
rules, manages parsing, invokes actions, tracks how often each rule matches
to optimise rule ordering (\sectionref{Rule ordering}), and stores results
in the database.  The framework is largely independent of the parser being
written: the same functionality will be required for most or all parsers,
though some specialised support functions may be provided.  The framework
is the core of the architecture and is deliberately quite simple: the rules
deal with the variation in inputs, and the actions deal with the
intricacies and complications encountered when parsing.

The function that finds the rule matching the input and invokes the
requested action can be expressed in pseudo-code as:

@@

Each action contains the code required to deal with a particular category
of input, e.g.\ rejections.  The actions are parser-specific: each parser
author will need to write the required actions from scratch, unless
extending an existing parser.  It is anticipated that parsers utilising
this architecture will have a high ratio of rules to actions, with the aim
of having simpler rules and clearer distinctions between the inputs parsed
by different rules.  In the Postfix log parser developed for this project
there are \numberOFactions{} actions and \numberOFrules{} rules, with an
uneven distribution of rules to actions as shown in \graphref{Distribution
of rules per action}.
\showgraph{build/graph-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} Unsurprisingly the action with
the most associated rules is \texttt{DELIVERY\_REJECTED}, the action that
handles Postfix rejecting a mail delivery attempt; it is followed by
\texttt{SAVE\_DATA}, the action responsible for handling informative log
lines, that supplement the data gathered from other log lines but do not by
themselves represent events.  The third most common action is, perhaps
surprisingly, \texttt{UNINTERESTING}: this action does nothing when
executed, allowing uninteresting log lines to be parsed without causing any
effects (it does not imply that the input is ungrammatical or unparsed).
Generally rules specifying the \texttt{UNINTERESTING} action are rules to
parse log lines that are not associated with a specific mail, e.g.\
notices about configuration files changing.  The remaining actions only
have one or two associated rules: in some cases this is because there will
only ever be one log line variant for that action, e.g.\ all log lines
showing that a remote client has connected are matched by a single rule and
handled by the \texttt{CONNECT} action; in other cases the action is
required to address a deficiency in the log files, or a complication that
arises during parsing.  Taking the \texttt{CONNECT} action as an example,
it creates a new data structure in memory for each client connecting, and
saves the data extracted by the rule; this data will subsequently be
entered into the database.  If a data structure already exists for that
connection it is treated as a symptom of a bug, and the action issues a
warning containing the full contents of the existing data structure, plus
the log line that has just been parsed.

The ability to add special purpose actions to deal with difficulties and
new requirements that are discovered during parser development is one of
the strengths of this architecture.  Instead of writing a single monolithic
function that must be modified to support new behaviour, with all the
attendant risks of adversely affecting the existing parser, when a new
requirement arises an independent action can be written to satisfy it.  
Sometime the new action will require the cooperation of other actions,
e.g.\ to set or check a flag.  There is a possibility of introducing
failure when modifying existing actions, but the modifications will be
smaller and occur less frequently than with a monolithic architecture, thus
failures will be less likely and will be easier to test for and diagnose.
The architecture can be implemented in an object oriented style, allowing
sub-classes to extend or override actions in addition to adding new
actions; because each action is an independent procedure, the sub-class
need only modify the action it is overriding, rather than reproducing large
chunks of functionality.

During development of the Postfix log parser it became apparent that in
addition to the obvious variety in log lines, there were many 
complications to be overcome.  Some were the result of deficiencies in
Postfix's logging (some of which were improved on or resolved by later
versions of Postfix); others were due to the vagaries of process
scheduling, client behaviour, and administrative actions.  All were
successfully accommodated in the Postfix log parser: adding new actions was
enough to overcome several of the complications; others required
modifications to a single existing action to work around the difficulties;
the remainder were resolved by adapting existing actions to cooperate and
exchange extra data, changing their behaviour as appropriate.

Actions may return a modified input line that will be parsed as if read
from the input stream, allowing for a simplified version of cascaded
parsing~\cite{cascaded-parsing}.  This is a powerful facility allowing 
several rules and actions to parse a single input.

@@

Rules categorise inputs, specifying both the regex to match against each
input and the action to invoke when the match is successful.  The Postfix
log parser stores the rules it uses in the same SQL database the results are stored
in, removing any doubt about which set of rules was used to produce a set
of results; other implementations are free to store the rules in whatever
fashion suits their needs.  The framework warns about every unparsed input,
to alert the user that they need to alter or extend their ruleset; the
Postfix log parser successfully parses every log line in the
\numberOFlogFILESall{} log files it is currently being tested with.  The
framework requires each rule to have \texttt{action} and \texttt{regex}
attributes; each implementation is free to add any additional attributes it
requires.  The Postfix log parser adds several attributes to optimise rule
ordering (\sectionref{Rule ordering}), to restrict which log lines each
rule can be matched against, and to describe each rule.  Data is
automatically extracted and saved from the log line based on keywords in
the rule's regex, but there is also a mechanism to specify extra data to be
saved.

Parsing new inputs is generally achieved by creating a new rule that pairs
an existing action with a new regex.  Adding new rules is simple: define
all the required attributes and append the new rule to wherever the
existing rules are stored.  The Postfix log parser supplies a utility based
on Simple Logfile Clustering Tool~\cite{slct-paper} to aid in producing
regexes from previously unparsed log lines.  Other approaches to rule
management can be used, e.g.\ instead of manually adding new rules, machine
learning techniques could be used to automatically generate new rules,
based on the existing rules and input lines.  If this approach was taken
the choice of machine learning technique would be constrained by the size
of typical data sets (see \sectionref{Results}).  Techniques that require
the full data set when training would be impractical: Instance Based
Learning~\cite{instance-based-learning} techniques that automatically
determine which inputs from the training set are valuable and which inputs
can be discarded might reduce the data required to a manageable
size.  This avenue of research and development has not been pursued by the
authors, but could easily be undertaken independently.

The architecture does not try to detect overlapping rules: that
responsibility is left to the author of the rules.  Unintentionally
overlapping rules lead to inconsistent parsing and data extraction because
the first matching rule wins, and the order in which rules are tried
against each input might change between parser invocations.  Overlapping
rules are frequently a requirement, allowing a more specific rule to match
some inputs and a more general rule to match the remainder, e.g.\
separating \SMTP{} delivery to specific sites from \SMTP{} delivery to the
rest of the world.  This can simplify both the general rule and the more
specific rule; additionally rules from different sources can be combined
with a minimum of prior cooperation or modification required.  Overlapping
rules should have a priority attribute to specify their relative ordering;
negative priorities may be useful for catchall rules.

Storing the rules separately to the parser allows external tools to be
written to detect overlapping rules.  Traditional regexes are equivalent in
computational power to \FA{}, and it is possible to convert between the
two, so regex overlap can be detected by finding a non-empty intersection
of two \FA{}\@.  The standard equation for \FA{} intersection (given for
example in~\cite{intersection-of-NFA-using-Z}) is: $FA1 \cap{} FA2 =
\overline{(\overline{FA1} \cup{} \overline{FA2})}$.  Perl 5.10 regexes are
more powerful than traditional regexes, e.g.\ it is possible to match
correctly balanced brackets nested to an arbitrary depth:
\newline{}\tab{}\verb!/^[^<>]*(<(?:(?>[^<>]+)|(?1))*>)[^<>]*$/!\newline{}
Perl 5.10 regexes can maintain an arbitrary state stack and are thus
equivalent in computational power to \PDA{} or \CFL{}, so detecting overlap
may require calculating the intersection of two \PDA{} or \CFL{}\@.  The
intersection of two \CFL{} is not closed, i.e.\ the resulting language
cannot in all cases be parsed by a \CFL{}, so intersection may be
intractable in some cases e.g.:
$a^{*}b^{n}c^{n}~\cap~a^{n}b^{n}c^{*}~\rightarrow~a^{n}b^{n}c^{n}$.

@@

It is possible to define pathological regexes that fall into two main
categories: regexes that match every input, and regexes that consume
excessive amounts of CPU time during matching.  Defining a regex that
matches every input is trivial: \verb!/^/! matches the start of every
input.  In general a regex will not consume much CPU time when the match is
successful; usually CPU time is consumed when a regex with a lot of
alteration and variable quantifiers fails to match.  This topic is beyond
the scope of this paper, see~\cite{mastering-regular-expressions} for
in-depth discussion.

@@

    \item [Partially context-sensitive actions and cascaded parsing:]
        Actions can consult the results (or lack of results) of previous
        actions during execution, providing some context sensitivity.
        Actions can not inspect past or future inputs, but actions can
        return a modified input that will be parsed as if read from the
        input stream, allowing for a simplified version of cascaded
        parsing~\cite{cascaded-parsing}.

@@

    \item [Closer to Natural Language Processing than using a fixed
        grammar:] Unlike traditional parsers such as those used when
        compiling a programming language, this architecture does not
        require a fixed grammar specification that inputs must adhere to.
        The architecture is capable of dealing with interleaved inputs, out
        of order inputs, and ambiguous inputs where heuristics must be
        applied --- all have arisen and been successfully accommodated in
        the Postfix log parser.

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important for this architecture because it is
expected that typical parsers will have a large number of rules.  There are
\numberOFrules{} rules in the full Postfix log parser ruleset, whereas the
minimum number of rules required to parse the \numberOFlogFILES{} log files
is \numberOFrulesMINIMUM{}, \numberOFrulesMINIMUMpercentage{} of the full
ruleset.  A second set of statistics was generated using the minimum
ruleset, and the parsing times compared to those generated using the full
ruleset: the percentage parsing time increase when using the full ruleset
instead of the minimal ruleset for optimal, shuffled and reversed orderings
is shown in \graphref{Percentage parsing time increase of maximum ruleset
over minimum ruleset}.  Clearly that the increased number of rules has
a noticeable performance impact with reverse ordering, and a lesser impact
with shuffled ordering.  The optimal ordering shows a mean increase of
\input{build/include-full-ruleset-vs-minimum-ruleset} in parsing time for a
\numberOFrulesMAXIMUMpercentage{} increase in the number of rules.  These
results show that the architecture scales very well as the number of rules
increases, and that optimally ordering the rules is an important
optimisation enabling this.
\showgraph{build/graph-full-ruleset-vs-minimum-ruleset}{Percentage parsing
time increase of maximum ruleset over minimum ruleset}{Percentage parsing
time increase of maximum ruleset over minimum ruleset}

@@

Correct parsing of log lines must be measured first.  Warnings are issued
for any log lines that are not parsed; no such warnings are issued while
parsing the \numberOFlogFILES{} log files, so it can be safely concluded
that there are zero false negatives.  False positives are harder to
quantify: manually verifying that the correct rule parsed each of the
\numberOFlogLINES{} log lines is infeasible.  A random sample of 6,039 log
lines (0.00994\% of \numberOFlogLINES{}) was parsed and the results
manually verified by inspection to ensure that the correct rule parsed each
log line.  The sample results contained zero false positives, and this
check has been automated to ensure continued accuracy.  The authors are
confident that zero false positives occur when parsing the
\numberOFlogFILES{} log files.

The proportion of mail delivery attempts correctly understood is much more
difficult to determine accurately than the proportion of log lines
correctly parsed.  The implementation can dump its state tables in a human
readable form; examining these tables with reference to the log files and
database is the best way to detect mails that were not handled properly.
The implementation issues warnings when it detects any errors or
discrepancies, which should alert the user to a problem.  There should be
few or no warnings during parsing, and when parsing is finished the state
table should only contain entries for mails starting before or ending after
the log file.  A second sample of 6000 log lines was parsed with all
debugging options enabled, resulting in 167,448 lines of output.  All
167,448 lines were examined in conjunction with the log segment and a dump
of the resulting database, verifying that for each of the log lines the
implementation performed correctly.  The implementation produced 4 warnings
about deficiencies in the log segment, 10 mails correctly remaining in the
state tables, and 1625 correct entries in the database.  No error messages
were produced, therefore there were no false negatives.  Given the evidence
detailed above, the authors are confident that the false positive rate when
understanding a mail delivery attempt is zero.

@@

This architecture's greatest strength is the ease with which it can be
adapted to deal with new requirements and inputs.  Parsing a variation of
an existing input is a trivial task: simply modify an existing rule or add
a new rule with an appropriate regex and the task is complete.  Parsing a
new category of input is achieved by writing a new action and appropriate
rules; quite often the new action will not need to interact with existing
actions, but when interaction is required the framework provides the
necessary facilities.  The decoupling of rules from actions allows
different sets of rules to be used with the same actions, e.g.\ a parser
might have actions to process versions one and two of a file format; the
parser will then parse version one, or version two, or both versions,
depending on the ruleset selected.  The architecture makes it possible to
apply commonly used programming techniques such as object orientation,
inheritance, composition, delegation, and roles when designing and
implementing a parser, simplifying the process of working within a team
with shared responsibility for a parser, or when developing and testing
additional functionality.  This architecture is ideally suited to parsing
inputs where the input is not fully understood or does not follow a fixed
grammar: the architecture warns about unparsed inputs and other errors, but
continues on parsing as best it can, allowing the developer of a new parser
to decide which deficiencies are most important and require attention
first, rather than being forced to fix the first error that arises.
